---
title: "Body Movement Predictions"
output: html_document
---

## Summary
This study attempted to classify movements based on movement data recorded through sensors worn by the subjects. Based on a Bagged Classification and Regression Tree Algorithm (CART) cross validated prediction accuracy was 98.1%.

## Data Processing

The data was a bit messy, hence the following processing was undertaken. There were no transformations applied because a subset of the data was used.

```{r, cache=TRUE}
# Load Data
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

# Process Training Data for Analysis
training$X <- NULL
training$user_name <- as.factor(training$user_name)
training$kurtosis_roll_belt <- as.numeric(training$kurtosis_roll_belt)
training$kurtosis_picth_belt <- as.numeric(training$kurtosis_picth_belt)
training$skewness_roll_belt <- as.numeric(training$skewness_roll_belt)
training$skewness_roll_belt.1 <- as.numeric(training$skewness_roll_belt.1)
training$max_yaw_belt <- as.numeric(training$max_yaw_belt )
training$min_yaw_belt <- as.numeric(training$min_yaw_belt)
training$amplitude_yaw_belt <- as.numeric(training$amplitude_yaw_belt)
training$kurtosis_roll_arm <- as.numeric(training$kurtosis_roll_arm)
training$kurtosis_picth_arm <- as.numeric(training$kurtosis_picth_arm)
training$kurtosis_yaw_arm <- as.numeric(training$kurtosis_yaw_arm)
training$skewness_roll_arm <- as.numeric(training$skewness_roll_arm)
training$skewness_pitch_arm <- as.numeric(training$skewness_pitch_arm)
training$skewness_yaw_arm <- as.numeric(training$skewness_yaw_arm)
training$kurtosis_yaw_arm <- as.numeric(training$kurtosis_yaw_arm)
training$kurtosis_roll_dumbbell <- as.numeric(training$kurtosis_roll_dumbbell)
training$kurtosis_picth_dumbbell <- as.numeric(training$kurtosis_picth_dumbbell)
training$skewness_roll_dumbbell <- as.numeric(training$skewness_roll_dumbbell)
training$skewness_pitch_dumbbell <- as.numeric(training$skewness_pitch_dumbbell)
training$kurtosis_yaw_belt <- as.numeric(training$kurtosis_yaw_belt)
training$skewness_yaw_belt <- as.numeric(training$skewness_yaw_belt)
training$kurtosis_yaw_dumbbell <- as.numeric(training$kurtosis_yaw_dumbbell)
training$skewness_yaw_dumbbell <- as.numeric(training$skewness_yaw_dumbbell)
training$max_yaw_dumbbell <- as.numeric(training$max_yaw_dumbbell)
training$min_yaw_dumbbell <- as.numeric(training$min_yaw_dumbbell)
training$amplitude_yaw_dumbbell <- as.numeric(training$amplitude_yaw_dumbbell)
training$kurtosis_roll_forearm <- as.numeric(training$kurtosis_roll_forearm)
training$kurtosis_picth_forearm <- as.numeric(training$kurtosis_picth_forearm)
training$kurtosis_yaw_forearm <- as.numeric(training$kurtosis_yaw_forearm)
training$skewness_roll_forearm <- as.numeric(training$skewness_roll_forearm)
training$skewness_pitch_forearm <- as.numeric(training$skewness_pitch_forearm)
training$skewness_yaw_forearm <- as.numeric(training$skewness_yaw_forearm)
training$max_yaw_forearm <- as.numeric(training$max_yaw_forearm)
training$min_yaw_forearm <- as.numeric(training$min_yaw_forearm)
training$amplitude_yaw_forearm <- as.numeric(training$amplitude_yaw_forearm)

# Apply the Same Processing to Test Data
testing$X <- NULL
testing$user_name <- as.factor(testing$user_name)
testing$kurtosis_roll_belt <- as.numeric(testing$kurtosis_roll_belt)
testing$kurtosis_picth_belt <- as.numeric(testing$kurtosis_picth_belt)
testing$skewness_roll_belt <- as.numeric(testing$skewness_roll_belt)
testing$skewness_roll_belt.1 <- as.numeric(testing$skewness_roll_belt.1)
testing$max_yaw_belt <- as.numeric(testing$max_yaw_belt )
testing$min_yaw_belt <- as.numeric(testing$min_yaw_belt)
testing$amplitude_yaw_belt <- as.numeric(testing$amplitude_yaw_belt)
testing$kurtosis_roll_arm <- as.numeric(testing$kurtosis_roll_arm)
testing$kurtosis_picth_arm <- as.numeric(testing$kurtosis_picth_arm)
testing$kurtosis_yaw_arm <- as.numeric(testing$kurtosis_yaw_arm)
testing$skewness_roll_arm <- as.numeric(testing$skewness_roll_arm)
testing$skewness_pitch_arm <- as.numeric(testing$skewness_pitch_arm)
testing$skewness_yaw_arm <- as.numeric(testing$skewness_yaw_arm)
testing$kurtosis_yaw_arm <- as.numeric(testing$kurtosis_yaw_arm)
testing$kurtosis_roll_dumbbell <- as.numeric(testing$kurtosis_roll_dumbbell)
testing$kurtosis_picth_dumbbell <- as.numeric(testing$kurtosis_picth_dumbbell)
testing$skewness_roll_dumbbell <- as.numeric(testing$skewness_roll_dumbbell)
testing$skewness_pitch_dumbbell <- as.numeric(testing$skewness_pitch_dumbbell)
testing$kurtosis_yaw_belt <- as.numeric(testing$kurtosis_yaw_belt)
testing$skewness_yaw_belt <- as.numeric(testing$skewness_yaw_belt)
testing$kurtosis_yaw_dumbbell <- as.numeric(testing$kurtosis_yaw_dumbbell)
testing$skewness_yaw_dumbbell <- as.numeric(testing$skewness_yaw_dumbbell)
testing$max_yaw_dumbbell <- as.numeric(testing$max_yaw_dumbbell)
testing$min_yaw_dumbbell <- as.numeric(testing$min_yaw_dumbbell)
testing$amplitude_yaw_dumbbell <- as.numeric(testing$amplitude_yaw_dumbbell)
testing$kurtosis_roll_forearm <- as.numeric(testing$kurtosis_roll_forearm)
testing$kurtosis_picth_forearm <- as.numeric(testing$kurtosis_picth_forearm)
testing$kurtosis_yaw_forearm <- as.numeric(testing$kurtosis_yaw_forearm)
testing$skewness_roll_forearm <- as.numeric(testing$skewness_roll_forearm)
testing$skewness_pitch_forearm <- as.numeric(testing$skewness_pitch_forearm)
testing$skewness_yaw_forearm <- as.numeric(testing$skewness_yaw_forearm)
testing$max_yaw_forearm <- as.numeric(testing$max_yaw_forearm)
testing$min_yaw_forearm <- as.numeric(testing$min_yaw_forearm)
testing$amplitude_yaw_forearm <- as.numeric(testing$amplitude_yaw_forearm)
```

## Data Exploration

A correlation matrix was calculated on the numeric variables.

```{r, cache=TRUE}
library(corrplot)
corrSubset <- subset(training[,sapply(training,is.numeric)], 
                     select = -c(kurtosis_yaw_belt, 
                                 skewness_yaw_belt, 
                                 kurtosis_yaw_dumbbell, 
                                 skewness_yaw_dumbbell, 
                                 kurtosis_yaw_forearm, 
                                 skewness_yaw_forearm))
correlation <- cor(corrSubset, use = "na.or.complete")
corrplot(correlation, order = "hclust", tl.pos = "n")
```

Due to the number of variables and considering the colinearity, the variables with greater than 5% correlation with the response variable were extracted:

```{r, cache=TRUE}
classeCor <- sapply(names(corrSubset), function(col) {
    cor(corrSubset[,col], as.numeric(training$classe))
})
classeCor <- classeCor[!is.na(classeCor)]
classeCor <- classeCor[abs(classeCor) >= 0.05]
classeCor <- classeCor[order(abs(classeCor), decreasing = TRUE)]
correlatedVars <- names(classeCor)
```

## Data Modeling

To compare the predictive capacity of the variables correlated to the response and including all the variables a Classification and Regression Tree (CART) model was built for each subset:

```{r, cache=TRUE}
# Data Modeling
library(caret)
library(doMC)
registerDoMC()
set.seed(1031)

controlObject <- trainControl(method = "cv", number = 10, verboseIter=TRUE)

## CART Model Using Highly Correlated Variables
cartModelCorVar <- train(x = subset(training, select = correlatedVars),
                         y = training$classe,
                         method = "rpart",
                         trControl = controlObject)
print(cartModelCorVar)

## CART Model Using All Variables
cartModel <- train(x = subset(training, select = -classe),
                   y = training$classe,
                   method = "rpart",
                   trControl = controlObject)
print(cartModel)
```

Including all the variables over the variables correlated to the response does not add significantly to the accuracy, hence only the latter set of variables was used.

The Naive Bayes and Bagged CART models were fitted to the dataset.

```{r, cache=TRUE}
## Naive Bayes Model Using Highly Correlated Variables
naiveBayesModel <- train(x = subset(training, select = correlatedVars),
                         y = training$classe,
                         method = "nb",
                         trControl = controlObject)
print(naiveBayesModel)

## Bagged CART Model Using Highly Correlated Variables
baggedCartModel <- train(x = subset(training, select = correlatedVars),
                         y = training$classe,
                         method = "treebag",
                         trControl = controlObject)
print(baggedCartModel)

# Data Prediction

predictions <- predict(baggedCartModel, newdata = testing)
```

The Bagged CART model has an accuracy of 98.1% compared to an accuracy of 65.4% for Naive Bayes using 10-fold cross validation.  

Finally, predictions were made using the Bagged CART model and found to be 100% accurate for the test cases.